Avoiding global deforestation by taxing land in agricultural production: The implications for global markets

### 2024-03-08 paper.Rmd
Update paper.Rmd

Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not contain certain behaviors. In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed. By combining the two models, we can fill each model missing behavior for a more complete description of the global impacts.

The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model. By solving the models together, we can expand the descriptive power of both.

@@

This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each hectare of land that is converted to agricultural land, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land. In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average hectare of farmland in 2022 at $1,537.8, the tax would equate to $1,081.1/hectare. In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $82.2 per hectare. The economic price of this would be transfered to consumers through lower agricultural output and higher market prices. Overall, this land tax would reduce consumer well-being by $92.1 billion or $10.60 per person in 2033. Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them. Thus, if policymakers wish to see deforestation stopped, they could instead use import tariffs on products produced on land converted from forestland for agricultural use.

>>>>

Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, and, instead, consider several important variables exogenous. In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed. By combining the two models, we can fill each model's missing behavior for a more complete description of the global impacts.

The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, which uses returns as an exogenous input, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model, which accepts change in available agricultural land as an exogenous input. Specifically, we  solve both models concurrently for variable `qo` for land in the GTAP model and the parameter `RENTA` in the GTM model, which are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model are exactly the same as the changes in rents and the associated changes in agricultural land (calculated as the non-forested area from the total available land). Solving both models together in this way therefore allows us to expand their descriptive power.

@@

This analysis shows that, at an estimated tax rate of 70.3 percent of the value of each hectare of land that is converted to agricultural land, global net deforestation would drop to roughly zero by 2033, preserving 144.2 million hectares of forestland that otherwise would have been converted to agricultural land. In the United States, as USDA-NASS (Land Values 2022 Summary (August 2022)) valued the average hectare of farmland in 2022 at $1,537.8, the tax would equate to $1,081.1/hectare. In Brazil, CEIC valued the average price of Brazilian land at roughly 2,500 Brazilian Reals, which would put the tax at roughly $82.2 per hectare. The economic price of this would be transfered to consumers through lower agricultural output and higher market prices. Overall, this land tax would reduce consumer well-being by $92.1 billion or $10.60 per person in 2033. Governments might be wary of farmer discontent were they to take such fiscal actions and thus might be reluctant to impose such taxes on them. Thus, if policymakers wish to see deforestation stopped, they could instead use import tariffs on products produced on land converted from forestland for agricultural use. Those scenarios, however, will be covered in a subsequent study.
